---
title: "P8106_group2recovery_primaryanalysis" 
author: "Yimin Chen (yc4195), Yang Yi (yy3307), Qingyue Zhuo (qz2493)"
output:
  pdf_document:
    toc: yes
    toc_depth: 2
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r, include = FALSE}
library(tidyverse)
library(caret)
library(corrplot)
library(ggplot2)
library(AppliedPredictiveModeling)
library(klaR)
library(MASS)
library(pROC)
library(rpart.plot)
library(randomForest)
library(ranger)
library(glmnet)
library(earth)
library(parallel)
library(doParallel)
library(mlbench)
library(pdp)
library(vip)
library(mgcv)
library(pls)
```

# Import and data manipulation

```{r, results='hide'}
# Load recovery.RData environment
load("./recovery.Rdata")

dat %>% na.omit()

# dat1 draw a random sample of 2000 participants Uni:3307
set.seed(3307)

dat1 = dat[sample(1:10000, 2000),]

dat1 = 
  dat1[, -1] %>% 
  mutate(
    gender = as.factor(gender),
    race = as.factor(race),
    smoking = as.factor(smoking),
    hypertension = as.factor(hypertension),
    diabetes = as.factor(diabetes),
    vaccine = as.factor(vaccine),
    severity = as.factor(severity),
    study = as.factor(
      case_when(study == "A" ~ 1, study == "B" ~ 2, study == "C" ~ 3)
      )
    )

# dat2 draw a random sample of 2000 participants Uni:2493
set.seed(2493)

dat2 = dat[sample(1:10000, 2000),]

dat2 = 
  dat2[, -1] %>% 
  mutate(
    gender = as.factor(gender),
    race = as.factor(race),
    smoking = as.factor(smoking),
    hypertension = as.factor(hypertension),
    diabetes = as.factor(diabetes),
    vaccine = as.factor(vaccine),
    severity = as.factor(severity),
    study = as.factor(
      case_when(study == "A" ~ 1, study == "B" ~ 2, study == "C" ~ 3)
      )
    )

# Merged dataset with unique observation
covid_dat = rbind(dat1, dat2) %>% 
  unique()

covid_dat2 = model.matrix(recovery_time ~ ., covid_dat)[, -1]

# Partition dataset into two parts: training data (70%) and test data (30%)
rowTrain = createDataPartition(y = covid_dat$recovery_time, p = 0.7, list = FALSE)

trainData = covid_dat[rowTrain, ]
testData = covid_dat[-rowTrain, ]

# matrix of predictors
x1 = covid_dat2[rowTrain,]
# vector of response
y1 = covid_dat$recovery_time[rowTrain]
# matrix of predictors
x2 = covid_dat2[-rowTrain,]
# vector of response
y2 = covid_dat$recovery_time[-rowTrain]

ctrl1 = trainControl(method = "repeatedcv", number = 10, repeats = 5)
```

# Model training

## Least squares
```{r}
num_cores <- detectCores() 
cl <- makePSOCKcluster(num_cores)
registerDoParallel(cl)

set.seed(2)
ls.fit = train(x1, y1,
               method = "lm",
               trControl = ctrl1)
summary(ls.fit)

stopCluster(cl)
registerDoSEQ()
```

## Ridge
```{r}
num_cores <- detectCores() 
cl <- makePSOCKcluster(num_cores)
registerDoParallel(cl)

set.seed(2)
ridge.fit = train(x1, y1,
                  method = "glmnet",
                  tuneGrid = expand.grid(alpha = 0,
                  lambda = exp(seq(6, -8, length=100))),
                  trControl = ctrl1
                  )

plot(ridge.fit, xTrans = log)

ridge.fit$bestTune

coef(ridge.fit$finalModel, s = ridge.fit$bestTune$lambda)

stopCluster(cl)
registerDoSEQ()
```

## LASSO
```{r}
num_cores <- detectCores() 
cl <- makePSOCKcluster(num_cores)
registerDoParallel(cl)

set.seed(2)
lasso.fit = train(x1, y1,
                  method = "glmnet",
                  tuneGrid = expand.grid(alpha = 1,
                  lambda = exp(seq(1, -6, length=100))),
                  trControl = ctrl1
                  )

plot(lasso.fit, xTrans = log)

lasso.fit$bestTune

coef(lasso.fit$finalModel, lasso.fit$bestTune$lambda)

stopCluster(cl)
registerDoSEQ()
```

## PCR & PLS tune
```{r}
# Find ncomp for PCR
set.seed(2)
pcr.mod = pcr(recovery_time ~ .,
              data = trainData,
              scale = TRUE,
              validation = "CV")

summary(pcr.mod)

pls.mod = plsr(recovery_time ~ .,
          data = trainData,
          scale = TRUE,
          validation = "CV")

summary(pls.mod)
```

The range of number of components considered is 1:18 for both PCR and PLS.

## PCR
```{r}
# PCR
set.seed(2)
ctrl2 = trainControl(method = "repeatedcv",
                     number = 10,
                     repeats = 5,
                     selectionFunction = "best")

pcr.fit = train(x1, y1,
                method = "pcr",
                tuneGrid = data.frame(ncomp = 1:18),
                trControl = ctrl2,
                preProcess = c("center", "scale"))

ggplot(pcr.fit, highlight = TRUE) + theme_bw()
```

## PLS
```{r}
set.seed(2)

pls.fit = train(x1, y1,
                method = "pls",
                tuneGrid = data.frame(ncomp = 1:18),
                trControl = ctrl2,
                preProcess = c("center", "scale"))

ggplot(pls.fit, highlight = TRUE)
```

## GAM
```{r}
num_cores <- detectCores() 
cl <- makePSOCKcluster(num_cores)
registerDoParallel(cl)

set.seed(2)
gam.fit = train(x = covid_dat2[rowTrain,],
                y = covid_dat$recovery_time[rowTrain],
                method = "gam",
                tuneGrid = data.frame(method = "GCV.Cp", select = c(TRUE,FALSE)),
                trControl = ctrl1)

gam.fit$bestTune
summary(gam.fit$finalModel)

stopCluster(cl)
registerDoSEQ()
```

## MARS
```{r}
num_cores <- detectCores() 
cl <- makePSOCKcluster(num_cores)
registerDoParallel(cl)

set.seed(2)

mars_grid = expand.grid(degree = 1:3,
                        nprune = 2:25)

mars.fit = train(x = covid_dat2[rowTrain,],
                 y = covid_dat$recovery_time[rowTrain],
                 method = "earth",
                 tuneGrid = mars_grid,
                 trControl = ctrl1)

ggplot(mars.fit)
mars.fit$bestTune
summary(mars.fit$finalModel)

stopCluster(cl)
registerDoSEQ()
```


## Regression tree
```{r}
num_cores <- detectCores() 
cl <- makePSOCKcluster(num_cores)
registerDoParallel(cl)

set.seed(2)

#Build a regression tree on the training data to predict the respons
rpart.fit = train(recovery_time ~ . ,
                  covid_dat[rowTrain,],
                  method = "rpart",
                  tuneGrid = data.frame(cp = exp(seq(-6,-2, length = 50))),
                  trControl =ctrl1)

rpart.fit$bestTune

#plot of the tree
ggplot(rpart.fit, highlight = TRUE)

rpart.plot(rpart.fit$finalModel)

stopCluster(cl)
registerDoSEQ()
```

## Random Forest
```{r}
num_cores <- detectCores() 
cl <- makePSOCKcluster(num_cores)
registerDoParallel(cl)

# Using Caret package to build random forest plot
rf.grid = expand.grid(mtry = 1:16,
                      splitrule = "variance",
                      min.node.size = 1:6)

set.seed(2)
rf.fit = train(recovery_time ~ .,
               covid_dat[rowTrain,],
               method = "ranger",
               tuneGrid = rf.grid,
               trControl = ctrl1)

# Best tunning parameter
rf.fit$bestTune

ggplot(rf.fit, highlight = TRUE)

stopCluster(cl)
registerDoSEQ()
```

## Boosting
```{r}
num_cores <- detectCores() 
cl <- makePSOCKcluster(num_cores)
registerDoParallel(cl)

gbm_grid = expand.grid(n.trees = c(100, 250, 500, 1000, 2000, 3000),
                       interaction.depth = 1:3,
                       shrinkage = c(0.0005,0.001,0.002),
                       n.minobsinnode = 1)
set.seed(2)
gbm.fit = train(recovery_time ~ . ,
                covid_dat[rowTrain,],
                tuneGrid = gbm_grid,
                trControl = ctrl1,
                method = "gbm",
                verbose = FALSE)

gbm.fit$bestTune
ggplot(gbm.fit, highlight = TRUE)

stopCluster(cl)
registerDoSEQ()
```

# Model selection

```{r}
resamp = resamples(list(ls = ls.fit,
                        ridge = ridge.fit,
                        lasso = lasso.fit,
                        pcr = pcr.fit,
                        pls = pls.fit,
                        gam = gam.fit,
                        mars = mars.fit,
                        rpart = rpart.fit,
                        rf = rf.fit,
                        boost = gbm.fit))

summary(resamp)
bwplot(resamp, metric = "RMSE")
```

From the summary table above we can observe that MARS model has the best performance (lowest RMSE), which indicates that it appears to be better fitted.

Then make prediction based on MARS model and calculate the test error.

```{r}
mars.fit$bestTune

summary(mars.fit$finalModel)
```

```{r}
set.seed(2)

mars.pred = predict(mars.fit, newdata = covid_dat2[-rowTrain,])

# Test error
mse.mars = mean((mars.pred - covid_dat$recovery_time[-rowTrain])^2)

mse.mars

rmse.mars = sqrt(mse.mars)
rmse.mars
```