---
title: "P8106_yiminchen_secondaryanalysis" 
author: "Yimin Chen (yc4195), Yang Yi (yy3307), Qingyue Zhuo (qz2493)"
output:
  pdf_document:
    toc: yes
    toc_depth: 2
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


```{r, echo = T, message = FALSE, results='hide', warning=FALSE,include = FALSE}
library(caret)
library(MASS)
library(mlbench)
library(pROC)
library(klaR)
library(tidyverse)
```
## Import and data manipulation

```{r, results='hide'}
# Load recovery.RData environment
load("./recovery.Rdata")

dat %>% na.omit()

# dat1 draw a random sample of 2000 participants Uni:3307
set.seed(3307)

dat1 = dat[sample(1:10000, 2000),]

dat1 = 
  dat1[, -1] %>% 
  mutate(
    recovery_time = as.factor(
      case_when(recovery_time <= 30 ~ "long", recovery_time > 30 ~ "short")
    ),
    gender = as.factor(gender),
    race = as.factor(race),
    smoking = as.factor(smoking),
    hypertension = as.factor(hypertension),
    diabetes = as.factor(diabetes),
    vaccine = as.factor(vaccine),
    severity = as.factor(severity),
    study = as.factor(
      case_when(study == "A" ~ 1, study == "B" ~ 2, study == "C" ~ 3)
      )
    )

# dat2 draw a random sample of 2000 participants Uni:2493
set.seed(2493)

dat2 = dat[sample(1:10000, 2000),]

dat2 = 
  dat2[, -1] %>% 
  mutate(
    recovery_time = as.factor(
      case_when(recovery_time <= 30 ~ "long", recovery_time > 30 ~ "short")
    ),
    gender = as.factor(gender),
    race = as.factor(race),
    smoking = as.factor(smoking),
    hypertension = as.factor(hypertension),
    diabetes = as.factor(diabetes),
    vaccine = as.factor(vaccine),
    severity = as.factor(severity),
    study = as.factor(
      case_when(study == "A" ~ 1, study == "B" ~ 2, study == "C" ~ 3)
      )
    )

# Merged dataset with unique observation
covid_dat = rbind(dat1, dat2) %>% 
  unique()

covid_dat2 = model.matrix(recovery_time ~ ., covid_dat)[, -1] #ignore intercept

# Partition dataset into two parts: training data (70%) and test data (30%)
rowTrain = createDataPartition(y = covid_dat$recovery_time, p = 0.7, list = FALSE)

trainData = covid_dat[rowTrain, ]
testData = covid_dat[-rowTrain, ]

# matrix of predictors
x1 = covid_dat2[rowTrain,]
# vector of response
y1 = covid_dat$recovery_time[rowTrain]
# matrix of predictors
x2 = covid_dat2[-rowTrain,]
# vector of response
y2 = covid_dat$recovery_time[-rowTrain]

ctrl1 = trainControl(method = "repeatedcv", number = 10, repeats = 5)
ctrl2 = trainControl(method = "cv",
                          classProbs = TRUE,
                          summaryFunction = twoClassSummary)
```

## Data visualization

## Model training

classification 
- classification tree: L11
- glm + penalized logistice regreesion L8
- GAM  L8
- MARS  L8
- QDA  L9
- LDA  L9
- Navie Bayes  L9
- random forest L12
- boosting L12
- support vecotr machines L13

## LDA
```{r}
set.seed(2)

model.lda <- train(x = covid_dat2[rowTrain,],
                   y = covid_dat$recovery_time[rowTrain],
                   method = "lda",
                   metric = "ROC",
                   trControl = ctrl2)
```

## QDA

```{r}
set.seed(2)
model.qda <- train(x = covid_dat2[rowTrain,],
                   y = covid_dat$recovery_time[rowTrain],
                   method = "qda",
                   metric = "ROC",
                   trControl = ctrl2)
```

## Naive Bayes (NB)

There is one practical issue with the NB classifier when nonparametric estimators are used. When a new data point includes a feature value that never occurs for some response class, the posterior probability can become zero. To avoid this, we increase the count of the value with a zero occurrence to a small value, so that the overall probability doesn't become zero. In practice, a value of one or two is a common choice. 
This correction is called "Laplace Correction," and is implemented via the parameter `fL`. The parameter `adjust` adjusts the bandwidths of the kernel density estimates, and a larger value means a more flexible estimate.

```{r, warning=FALSE}
nbGrid <- expand.grid(usekernel = c(FALSE,TRUE),
                      fL = 1, 
                      adjust = seq(.2, 3, by = .2))

set.seed(2)
model.nb <- train(x = covid_dat2[rowTrain,],
                  y = covid_dat$recovery_time[rowTrain],
                  method = "nb",
                  tuneGrid = nbGrid,
                  metric = "ROC",
                  trControl = ctrl2)

plot(model.nb)
```

```{r}
res <- resamples(list(LDA = model.lda, QDA = model.qda, NB = model.nb))
summary(res)
```

# test set performance.
```{r}
lda.pred <- predict(model.lda, newdata = covid_dat2[-rowTrain,], type = "prob")[,2]
nb.pred <- predict(model.nb, newdata = covid_dat2[-rowTrain,], type = "prob")[,2]
qda.pred <- predict(model.qda, newdata = covid_dat2[-rowTrain,], type = "prob")[,2]


roc.lda <- roc(covid_dat$recovery_time[-rowTrain], lda.pred)
roc.nb <- roc(covid_dat$recovery_time[-rowTrain], nb.pred)
roc.qda <- roc(covid_dat$recovery_time[-rowTrain], qda.pred)


auc <- c(roc.lda$auc[1], roc.qda$auc[1], roc.nb$auc[1])

plot(roc.lda, legacy.axes = TRUE)
plot(roc.qda, col = 2, add = TRUE)
plot(roc.nb, col = 3, add = TRUE)

modelNames <- c("lda","qda","nb")
legend("bottomright", legend = paste0(modelNames, ": ", round(auc,3)),
       col = 1:3, lwd = 2)
```
